---
title: "Gov 50 Final Project"
author: "Allison Hunter"
description: "Gender and Race Dynamics and Allegations of Crime in the Era of Lynchings in 19th Century America"
output:
  distill::distill_article:
    self_contained: false
---

## Research Question and Interest

I became interest in the climate change conversation as Harvard was my true  introduction to the topic due to my school not prioritizing the education of this issue (I went to a low-income public high school and our curriculum was truly all over the place). I had no clue about the impact of carbon emissions and how they contributed to the warming of the Earth and wanted to do research to see how it impacted various parts of our planet. 

My research question is "How does Carbon Emissions affect global temperatures 
of the atmosphere, land and ocean?". As carbon emissions increase, I hypothesize that there will be a corresponding rise in temperatures across the atmosphere, land, and oceans. This hypothesis is grounded in the well-established relationship between increased levels of carbon dioxide (CO2) leading to heat retention in the Earth's atmosphere. This phenomenon is expected to contribute to elevated temperatures in the oceans, atmosphere, and on land." The explanatory variable in this research would be the carbon emissions and the outcome variable the average temperature of the ocean, land, and atmosphere. The observed pattern in my data would prove my hypothesis correct if we see a increase in temperatures as carbon emissions increase. My hypothesis would be disapproved if we saw a decrease in temperatures if carbon emissions increase.

## Data

My original data was taken from 



```{r}

library(tidyverse)
carbon <- read.csv("temp_carbon.csv")
carbon

```

```{r}
library(ggplot2)
ggplot(data = carbon,
       mapping = aes(x = temp_anomaly,
                      y = carbon_emissions, 
                     color = year)) +
  geom_point() + geom_smooth() + 
  
  labs(x = "Global Average Temperature in Celsius",
       y = "Annual Carbon Emissions in Millions of Metric Tons",
       title = "Global Average Temp vs Annual Carbon Emissions")

ggplot(data = carbon,
       mapping = aes(x = land_anomaly,
                      y = carbon_emissions, 
                     color = year)) +
  geom_point() + geom_smooth() +
   labs(x = "Average Temperature of Land in Celsius",
       y = "Annual Carbon Emissions in Millions of Metric Tons",
       title = "Average Land Temperature vs Annual Carbon Emissions")

ggplot(data = carbon,
       mapping = aes(x = ocean_anomaly,
                      y = carbon_emissions, 
                     color = year)) +
  geom_point() + geom_smooth() +
  labs(x = "Average Temperature of Ocean in Celsius",
       y = "Annual Carbon Emissions in Millions of Metric Tons",
       title = "Average Ocean Temperature vs Annual Carbon Emissions")

```

```{r}
carbon_fit <- lm(carbon_emissions ~ land_anomaly + temp_anomaly + ocean_anomaly,
               data = carbon) 
summary(carbon_fit)

var_labels <- c(
"(Intercept)" = "Intercept",
"land_anomaly" = "Avg Land Temp",
"temp_anomaly" = "Global Avg Temp",
"ocean_anomaly" = "Avg Ocean Temp"
)

modelsummary::modelsummary(carbon_fit,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"), 
                           coef_map = var_labels,
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```

